Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-22 paper.tex
reply
bisim S D

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation.  It is
the smallest congruence relation that also includes the pairs induced
by the rules in Fig.~\ref{fig:bisim-SC-D}.
%
The relation is mutually defined on continuations, written $\kappa
\approx \kappa$, and values, written $v \approx v : T$, where $T$ is
their type.
%
If $k \approx \kappa$, adding an identity cast on top of $k$ gives a related 
continuation.
%
\todo[inline]{The use of ``wrap'' obscures this explanation.
  Consider getting rid of wrap by inlining it. -JS}
%
If $\kappa_1 \approx \kappa_2$, extending $\kappa_1$ and $\kappa_2$ with the 
same cast gives related results.
%
If two values of type $P$ are related, injecting them to $\TOOdyn$ gives 
related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a non-casted function in \ineffCEKD, the function in \effCEK{C} 
must use identities as its casts.
%
If functions are related, we can apply the same cast on both side.
%
Rules for products and sums are similar.

@@

\begin{lemma}[SoAndSo]
	Assume $C$ implements Lazy D Cast ADT
	and $v_1$ is a value in \effCEK{C}
	and $v_2$ is a value in \ineffCEKD\
	and $v_1 \approx v_2 : T_1$,
	\[
	applyCast_{\text{\effCEK{C}}}(v,cast(T_1,l,T_2)) 
	\approx 
	applyCast_{\text{\ineffCEKD}}(v,\cOOcast{T_1}{l}{T_2})
	\]
\end{lemma}
\begin{proof}
	Case splitting whether $T_1$ is shallowly consistent with $T_2$.
	If yes, case splitting how they are shallowly consistent.
	Then apply properties (3) - (11) of Lazy D Cast when applicable.
\end{proof}

@@

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
	\label{thm:surely-monoidic-reduce}
	Assume $C$ implements Lazy D Cast ADT
	and $C$ is a monoid
	and $cast(T,l,T)=id(T)$,
	$ s_{1} \in \text{\effCEK{C}}$,
	$ s_{2} \in \text{\ineffCEKD}$,
	$ s_{1} \approx s_{2}$,
\begin{enumerate}
	\item If $s_1 = \sOOhalt{o}$ then $s_2 = \sOOhalt{o}$
	\item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
	\item If $\judgeSreduce{C}{s_1}{s_3}$ 
		then $\judgeDreduce{s_2}{s_4}$ for some $s_4$
	\item If $\judgeDreduce{s_2}{s_4}$
		then $\judgeSreduce{C}{s_1}{s_3}$ for some $s_3$
	\item If $\judgeSreduce{C}{s_1}{s_3}$ and $\judgeDreduce{s_2}{s_4}$
		then $s_3 \approx^{*} s_4$
\end{enumerate}
%	Or diagrammatically, 
%	\[\begin{array}{clclc}
%	s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
%	\rotatebox[origin=c]{90}{$\approx$} 
%	& & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
%	s_{2} & 
%	\longrightarrow_{\text{\ineffCEKD}} & s_2' &
%	\longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
%	\end{array}\]

%	$ s_{1}, s_{1}' \in \text{\effCEK{C}}$,
%	$ s_{2}, s_{2}' \in \text{\ineffCEKD}$,
%	$ s_{1} \approx s_{2}$,	
%	$ \judgeSreduce{C}{s_{1}}{s_{1}'}$, and
%	$ \judgeDreduce{s_{2}}{s_{2}'}$, then
%	\[ \exists s_2'' \in \text{\ineffCEKD}: \;
%	\judgeDreduceTrans{s_{2}'}{s_{2}''} \;\text{and}\; s_1' \approx s_2'' \]
%	Or diagrammatically, 
%	\[\begin{array}{clclc}
%	s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
%	\rotatebox[origin=c]{90}{$\approx$} 
%	& & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
%	s_{2} & 
%	\longrightarrow_{\text{\ineffCEKD}} & s_2' &
%	\longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
%	\end{array}\]

\end{lemma}
\begin{proof}
	Part (1) and (2) follow from the definition of $s \approx s$.
	Part (3) and (4) are true because we have transition rules for every 
	non-halting state.
	Part (5) is more interesting.
	Consider
	\[
	\sOOreturn{\hcvOOcons{v_1}{c_1}{v_2}{c_2}}{\kOOcast{c}{\kOOcar{\kappa}}}
	\approx
	\]
\end{proof}

\todo[inline]{UNDER CONSTRUCTION -KC}

>>>>

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation.  It is
the smallest congruence relation that also includes the pairs induced
by the rules in Fig.~\ref{fig:bisim-SC-D}.
%
The relation is mutually defined on continuations, written $\kappa
\approx \kappa$, and values, written $v \approx v : T$, where $T$ is
their type.
%
If $k \approx \kappa$, adding an identity cast on top of $k$ gives a related 
continuation.
%
\todo[inline]{The use of ``wrap'' obscures this explanation.
  Consider getting rid of wrap by inlining it. -JS
\\---\\
I agree that inlining wrap and ext can improve readability of $\kappa \approx 
\kappa$, but this would make latex different from Agda, is it alright?
-KC}
%
If $\kappa_1 \approx \kappa_2$, extending $\kappa_1$ and $\kappa_2$ with the 
same cast gives related results.
%
If two values of type $P$ are related, injecting them to $\TOOdyn$ gives 
related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a non-casted function in \ineffCEKD, the function in \effCEK{C} 
must use identities as its casts.
%
If functions are related, we can apply the same cast on both side.
%
Rules for products and sums are similar.

@@

\begin{lemma}[SoAndSo]
	\label{thm:lem-apply-cast-SD}
	Assume $C$ implements Lazy D Cast ADT
	and $v_1$ is a value in \effCEK{C}
	and $v_2$ is a value in \ineffCEKD\
	and $v_1 \approx v_2 : T_1$,
	\[
	applyCast_{\text{\effCEK{C}}}(v,cast(T_1,l,T_2)) 
	\approx 
	applyCast_{\text{\ineffCEKD}}(v,\cOOcast{T_1}{l}{T_2})
	\]
\end{lemma}
\begin{proof}
	Case splitting whether $T_1$ is shallowly consistent with $T_2$.
	If yes, case splitting how they are shallowly consistent.
	Then apply properties (3) - (11) of Lazy D Cast when applicable.
\end{proof}

@@

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
	\label{thm:surely-monoidic-reduce}
	Assume $C$ implements Lazy D Cast ADT
	and $C$ is a monoid
	and $cast(T,l,T)=id(T)$,
	$ s_{1} \in \text{\effCEK{C}}$,
	$ s_{2} \in \text{\ineffCEKD}$,
	$ s_{1} \approx s_{2}$,
\begin{enumerate}
	\item If $s_1 = \sOOhalt{o}$ then $s_2 = \sOOhalt{o}$
	\item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
	\item If $\judgeSreduce{C}{s_1}{s_3}$ 
		then $\judgeDreduce{s_2}{s_4}$ for some $s_4$
	\item If $\judgeDreduce{s_2}{s_4}$
		then $\judgeSreduce{C}{s_1}{s_3}$ for some $s_3$
	\item If $\judgeSreduce{C}{s_1}{s_3}$ 
		and  $\judgeDreduce{s_2}{s_4}$ 
		then $s_3 \approx^{*} s_4$
\end{enumerate}
	Or diagrammatically, 
	\[\begin{array}{clclc}
	s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_1'\\
	\rotatebox[origin=c]{90}{$\approx$} 
	& & & \;\;\;\rotatebox[origin=c]{-20}{$\approx$} \\
	s_{2} & 
	\longrightarrow_{\text{\ineffCEKD}} & s_2' &
	\longrightarrow^{*}_{\text{\ineffCEKD}} & s_2'' \\
	\end{array}\]
\end{lemma}
\begin{proof}
	Part (1) and (2) follow from the definition of $s \approx s$.
	Part (3) and (4) are true because we have transition rules for every 
	non-halting state.
	Part (5) is more interesting. Our goal is to prove that one reduction step 
	in 	\effCEK{C} corresponds to one or more steps in \ineffCEKD.	
	Consider
	\[
	\sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{
		wrap(\kOOcar{\kappa})
	}
	\approx
	\sOOreturn{v_2}{\kOOcar{\kappa}}
	\]

	By looking at the definition of $v \approx v$, we know
	\[ c_{11} = seq(\dots seq(id(T_1),cast(T_1,l,T_3)) \dots) \]
	\[ c_{12} = seq(\dots seq(id(T_2),cast(T_2,l,T_4)) \dots) \] 
	\[ v_2 = \vOOcast{(\vOOcons{v_{21}}{v_{22}})}{
		\cOOcast{\POOprod{T_1}{T_2}}{l_1}{\POOprod{T_3}{T_4}}} \dots
	\]
	Thus we have the following reduction in \effCEK{C}. The equation 
	reasoning employs the fact that $C$ is a monoid.
	\[
	\begin{array}{ll}
&
\sOOreturn{\hcvOOcons{v_{11}}{c_{11}}{v_{12}}{c_{12}}}{wrap(\kOOcar{\kappa})}
\\
\longmapsto_\text{\effCEK{C}} &
\sOOreturn{v_{11}}{ext(c_{11},\kappa)}
\\
= &
\sOOreturn{v_{11}}{ext(seq(\dots seq(id(T_1),cast(T_1,l,T_3))\dots),\kappa)}
\\
= &
\sOOreturn{v_{11}}{ext(cast(T_1,l,T_3),\dots,\kappa)}
	\end{array}
	\]

And we have the following reduction in \ineffCEKD, where the last state is 
related to the last state in the above sequence.

\[
\begin{array}{ll}
& \sOOreturn{
	\vOOcast{(\vOOcons{v_{21}}{v_{22}})}{		
	\cOOcast{\POOprod{T_1}{T_2}}{l_1}{\POOprod{T_3}{T_4}}} \dots
}{\kOOcar{\kappa}}
\\
\longmapsto_{\mathcal{D}}^{*} &
\sOOreturn{(\vOOcons{v_{21}}{v_{22}})}{
	\kOOcar{
	\kOOcast{\cOOcast{T_1}{l}{T_3}}{\dots\kappa}
}}
\\
\longmapsto_{\mathcal{D}} &
\sOOreturn{v_{21}}{
	\kOOcast{\cOOcast{T_1}{l}{T_3}}{\dots\kappa}
}
\end{array}
\]

Other cases are similar. Of course we need the lemma about $applyCast$ 
(\ref{thm:lem-apply-cast-SD}) when cast applications happen in both machines. 
And in case spliting, we shall use $cast(T,l,T)=id(T)$ to relate the 
continuation functions.
\end{proof}

\todo[inline]{Should I invest time on restating the bisimulation lemma in the 
standard format? -KC}
